### 2022-08-15 readme.Rmd
python fig draft added

---
title: "Reproducible and Programmatic Generation of Neuroimaging Visualizations"
output: github_document
---

## Reference

c.	Chopra, S., Labache, L., Dhamala, E., Orchard, E., Holmes. A. [Reproducible and programmatic generation of neuroimaging visualizatio](https://github.com/sidchop/RepoNeuroVis/blob/main/manuscript.pdf). Preprint.

------------------------------------------------------------------------

## Background

Neuroimaging visualization forms the centrepiece of quality control, and interpretation and communication of results. Often, these images and figures are produced by manually changing settings on Graphical User Interfaces (GUI). There now exist many well-documented code-based brain visualisation tools that allow users to programmatically generate publication-ready figures directly within environments such as R, Python and MATLAB. Compared to figures generated using GUIs, programmatic tools are more replicable, flexible, interactive, and integrated with the scientific process. This article reviews the advantages of learning and using programmatic neuroimaging visualization tools and provides a list of tools across programming environments. Finally, we introduce examples of specific R and Python tools for visualization of voxel, vertex, region-of-interest, and edge level data..

------------------------------------------------------------------------

## Figures made in R: \

```{r, echo = FALSE,warning=F, error=F, out.width='100%'}
knitr::include_graphics("data/readme_figures/fig1.png")
```

## Figures made in Python: \

```{r, echo = FALSE,warning=F, error=F, out.width='100%'}
knitr::include_graphics("data/readme_figures/fig2.png")
```

------------------------------------------------------------------------

## Questions

Please contact me (Sidhant Chopra) as [sidhant.chopra\@yale.edu](mailto:sidhant.chopra@yale.edu){.email} and/or [sidhant.chopra4\@gmail.com](mailto:sidhant.chopra4@gmail.com){.email}