### 2022-08-15 readme.Rmd
tabel added to rdme

---
title: "Reproducible and Programmatic Generation of Neuroimaging Visualizations"
output: github_document
---

## Reference

**Chopra, S.**, Labache, L., Dhamala, E., Orchard, E., Holmes. A. [Reproducible and programmatic generation of neuroimaging visualizatio](https://github.com/sidchop/RepoNeuroVis/blob/main/manuscript.pdf). Preprint.

------------------------------------------------------------------------

## Background

Neuroimaging visualization forms the centrepiece of quality control, and interpretation and communication of results. Often, these images and figures are produced by manually changing settings on Graphical User Interfaces (GUI). There now exist many well-documented code-based brain visualisation tools that allow users to programmatically generate publication-ready figures directly within environments such as R, Python and MATLAB. Compared to figures generated using GUIs, programmatic tools are more replicable, flexible, interactive, and integrated with the scientific process. This article reviews the advantages of learning and using programmatic neuroimaging visualization tools and provides a list of tools across programming environments. Finally, we introduce examples of specific R and Python tools for visualization of voxel, vertex, region-of-interest, and edge level data..

------------------------------------------------------------------------

## Figures made in R: \

```{r, echo = FALSE,warning=F, error=F, out.width='100%'}
knitr::include_graphics("data/readme_figures/fig1.png")
```

## Figures made in Python: \

```{r, echo = FALSE,warning=F, error=F, out.width='100%'}
knitr::include_graphics("data/readme_figures/fig2.png")
```

------------------------------------------------------------------------

## Examples of code-based neuroimaging visualizations tools that can be accessed directly within R, MATLAB and Python environments.

|                                                        | Voxel   | Vertex  | ROI           | Edge    | Streamlines |
|:-----------|:-----------|:-----------|:-----------|:-----------|:-----------|
| **R**                                                  |         |         |               |         |             |
| [ANTsR](https://github.com/ANTsX/ANTsR)                | ***+*** | ***+*** | ***+***       |         |             |
| [brainconn](https://github.com/sidchop/brainconn)      |         |         |               | ***+*** |             |
| [brainR](https://github.com/muschellij2/brainR)        | ***+*** |         | ***+***       |         |             |
| [ciftitools](https://github.com/mandymejia/ciftiTools) | ***+*** | ***+*** | ***+*****\*** |         |             |
| [fsbrain](https://github.com/dfsp-spirit/fsbrain)      | ***+*** | ***+*** | ***+*****\*** |         |             |
| ggseg                                                  |         |         | ***+***       |         |             |
| neurobase                                              | ***+*** |         |               |         |             |
| oro.nifti                                              | ***+*** |         |               |         |             |
| **Python**                                             |         |         |               |         |             |
| ANTsPy                                                 | ***+*** | ***+*** | ***+***       |         |             |
| brainiak                                               | ***+*** |         |               |         |             |
| Brainplotlib                                           |         | ***+*** | ***+*****\*** |         |             |
| Brainspace/surfplot                                    |         | ***+*** | ***+*****\*** |         |             |
| DIPY                                                   | ***+*** |         |               |         | ***+***     |
| ENIGMA toolbox                                         |         |         | ***+***       |         |             |
| FSLeyes                                                | ***+*** | ***+*** | ***+***       |         |             |
| ggseg                                                  |         |         | ***+***       |         |             |
| graphpype                                              |         |         |               | ***+*** |             |
| MMVT                                                   |         | ***+*** | ***+***       | ***+*** |             |
| MNE                                                    | ***+*** | ***+*** | ***+***       |         |             |
| mrivis                                                 | ***+*** |         |               |         |             |
| NaNSlice                                               | ***+*** |         |               |         |             |
| netneurotools                                          |         | ***+*** | ***+*****\*** |         |             |
| nilearn                                                | ***+*** | ***+*** | ***+***       | ***+*** |             |
| niwidget                                               | ***+*** | ***+*** |               |         | ***+***     |
| Pycortex                                               | ***+*** | ***+*** | ***+*****\*** |         |             |
| pySurfer                                               |         | ***+*** | ***+*****\*** |         |             |
| surfice                                                | ***+*** | ***+*** | ***+***       | ***+*** | ***+***     |
| surfplot                                               |         | ***+*** | ***+***       |         |             |
| Visbrain                                               | ***+*** | ***+*** | ***+***       | ***+*** |             |
| **MATLAB**                                             |         |         |               |         |             |
| BrainNetViewer                                         | ***+*** |         | ***+***       | ***+*** |             |
| Brainspace                                             |         | ***+*** | ***+*****\*** |         |             |
| Brainstorm                                             | ***+*** | ***+*** | ***+***       |         |             |
| bspmview                                               | ***+*** |         | ***+***       |         |             |
| CandlabCore                                            | ***+*** |         | ***+***       |         |             |
| ECoG/fMRI Vis toolbox                                  |         | ***+*** | ***+*****\*** |         |             |
| ENIGMA toolbox                                         |         |         | ***+***       |         |             |
| FieldTrip                                              | ***+*** | ***+*** |               |         |             |
| Lead-DBS                                               | ***+*** |         | ***+***       |         |             |
| mni2fs                                                 |         | ***+*** |               |         |             |
| mrtools                                                | ***+*** | ***+*** | ***+*****\*** |         |             |
| plotSurfaceROIBoundary                                 |         | ***+*** | ***+*****\*** |         |             |
| Vistasoft                                              | ***+*** | ***+*** | ***+***       |         | ***+***     |

Note: The tools listed contain functionality required to generate (at least close-to) publication-ready neuroimaging figures via user-entered code within R, MATLAB and Python environments. This list does not include cross-platform general purpose visualization software.\
\*Cortex only

------------------------------------------------------------------------

## Questions

Please contact me (Sidhant Chopra) as [sidhant.chopra\@yale.edu](mailto:sidhant.chopra@yale.edu){.email} and/or [sidhant.chopra4\@gmail.com](mailto:sidhant.chopra4@gmail.com){.email}