### 2022-08-31 SupplementA_R_Figs.Rmd
sup sep

---
title: "Supplement 1 - code for R figures"
output: html_notebook
---

```{r Fig1A, eval=FALSE}
#Figure 1A (Voxel)

library(neurobase)

#Load in a nifti (.nii.gz) file of a standardized template so use as a background. 
template <- readnii('data/MNI152_T1_1mm_brain.nii.gz') #load the nifti file into R

#Load in a nifti (.nii.gz) file of the statistic you wanted of overlay, with the same dimensions as the template.  
effect <- readnii("data/MNI152_effect_size.nii.gz")

#Threshold the statistic map so only values about a specified amount are displayed
effect[effect<0.3] <- NA

#Set the breaks/intervals you want on the color bar (e.g. from .3 to .6, by intervals of 0.005)
breaks = seq(.3,.6, by=0.005)

#[Optional] If you want to output the figure as a .png, open a png image device
png("data/R/fig1a.png")

#Use the ortho2 function from the neurobase package to make the figure
ortho2(x = template,   #Specify the background template
       y = effect,     #Specify the effect you want to overlay on the template
       crosshairs = F, #Remove the cross-hairs
       bg = "white",   #Make the background white 
       NA.x = T,       #Do not display NA values
       col.y= rev(viridis::inferno(n = 500)), #Select the color scale. 
       xyz = c(70,50,80), #set the x y & z slice you want to visualize
       useRaster = T,     #Sometimes using Raster makes for clearer plots
       ycolorbar = TRUE,  #Add a color bar
       mfrow = c(1,1)) +  #layout of brain views (for this figure, we just want a axial view)
  colorbar(breaks, col = rev(viridis::inferno(n=length(breaks)-1)), #add a color bar with same color scale 
           text.col = "black", labels = FALSE, maxleft = 0.85)  #Set the specifications for the color bar

dev.off()

```

```{r Fig1B, eval=FALSE}
#Figure 1B (Voxel)

library(neurobase)
library(scales)

#Load in a nifti (.nii.gz) file to use as a background. 
t1 <- readnii('data/sub-001__t1_warped.nii.gz')

#Load in a nifti (.nii.gz) file to use as a overlay. 
atlas <-  readnii('data/sub-001_schaefer300n7_aseg_to_dwispace_gm_rois.nii.gz')

png('data/R/fig1b.png')
overlay(x=robust_window(t1), y=atlas, 
        plot.type = "single", #Single view (axial only)
        z=c(seq(30,110,10)), #set slices you want to view (30 to 110, with a gap of 10 slices). 
        col.y =  alpha(viridis::inferno(300), 0.6),
        plane = "axial",  
        useRaster = T, 
        bg = "white",
        NA.x=T, #Make NA values on overlay transparent
        zlim.y =c(1,300))
dev.off()

```

```{r Fig1C, eval=FALSE}
#Figure 1C (Cortical ROI using ciftiTools)

# Load the package 
library(ciftiTools)
library(rgl)

#Need connectome workbench to use ciftiTools - set path below
ciftiTools.setOption("wb_path", "/Applications/workbench/")

#Define CIFTI file names for time series and surface files. 
cifti_fname <- ciftiTools::ciftiTools.files()$cifti["dtseries"]
surfL_fname <- ciftiTools.files()$surf["left"]
surfR_fname <- ciftiTools.files()$surf["right"]

# Read in CIFTI file
cii <- read_cifti(
  cifti_fname, brainstructures="all", 
  surfL_fname=surfL_fname,
  surfR_fname=surfR_fname)

set.seed(1993) #set a random seed (good practice for reproducibility)

parc <- load_parc("Schaefer_400") #ciftitools has a few atlas in cifti format already included

ramdom_metric <- rnorm(400) #generate random data to plot on cortex

cii <- move_from_mwall(cii, NA)

parc_vec <- c(as.matrix(parc))

xii_metric <- c(NA, ramdom_metric)[parc_vec + 1]

xii1 <- select_xifti(cii, 1)

xii_metric <- newdata_xifti(xii1, xii_metric)

view_xifti_surface(xii_metric, 
                   colors = viridis::viridis(n = 400), 
                   borders = "black", 
                   hemisphere = "both",
                   view = 'lateral')

rgl.snapshot("data/R/fig1c.png")

rgl.close()

```

```{r Fig1D, eval=FALSE}
#Figure 1D (Subcortical ROI using ggseg)

library(ggseg)
library(ggplot2)

sub_base_atlas <- as.data.frame(na.omit(cbind(aseg$data$region, aseg$data$hemi)))

colnames(sub_base_atlas) <- c("region", "hemi")

Effect <- rnorm(dim(sub_base_atlas)[1])

sub_base_atlas <- cbind(Effect, sub_base_atlas)
subcortex <- ggseg(atlas = aseg,
                   .data = sub_base_atlas,
                   mapping=aes(fill=Effect), 
                   position = "dispersed",
                   colour="black",
                   hemi =c('left', 'right'),
                   size=.2,
                   show.legend = T, 
                   plot.background = "white") + 
  scale_fill_viridis_b() +
  theme(legend.text=element_text(size=15),
        legend.title  = element_blank())

ggsave(filename = 'data/R/fig1d.png', plot = subcortex, device = 'png', bg = 'white')
```

```{r Fig1E, eval=FALSE}
#Figure 1E (Subcortical ROI using ggseg3D)

#The function below plots 3d version of aseg. See "inputs" within function for more details. 
make_ggseg3d <- function(attribute, 
                         colour.pal = c("light yellow",
                                        "orange",
                                        "red",
                                        "dark red"), 
                         hide.colourbar=FALSE, 
                         output.png=FALSE, 
                         file.name="ggseg_3d.png") {
  # Inputs: 
  #' attribute = This must be a numeric vector in this: Left-Thalamus, Left-Caudate, Left
  #' -Putamen, Left-Pallidum Left-Hippocampus, Left-Amygdala, Left-Accumbens-area, Righ
  #' t-Thalamus-Proper, Right-Caudate, Right-Putamen, Right-Pallidum, Right-Hippocampus, Righ
  #' t-Amygdala, Right-Accumbens-area

  library(ggseg3d)
  library(tidyr)
  library(dplyr)
  #remove(aseg_3d)
  aseg_3d <- aseg_3d
  aseg_3d <- tidyr::unnest(aseg_3d, cols = c(ggseg_3d))

  attribute.ggseg3d <- c(rep(NA, 4), attribute[1:4], rep(NA, 3), attribute[5:7],
                         rep(NA, 5), attribute[8:14],  rep(NA, 6))

  data <- dplyr::mutate(aseg_3d, attribute =  attribute.ggseg3d)
  #remove NA regions
  aseg_3d[which(is.na(data$attribute)),] <- NA
  aseg_3d <- tidyr::drop_na(aseg_3d)

  data[which(is.na(data$attribute)),] <- NA
  data <- tidyr::drop_na(data)
  data$attribute[data$attribute==0]<-NA #make 0 valus NA so they are set as grey in ggseg3d

  scene=list(camera = list(eye = list(x = 0, y = 1, z = -2.25)),
             aspectratio = list(x=1.6,y=1.6,z=1.6))

  plot <- ggseg3d::ggseg3d(.data = data, 
                           atlas = aseg_3d, 
                           colour = "attribute", 
                           text = "attribute",
                           palette = colour.pal)
  plot <- remove_axes(plot)
  plot <- plotly::layout(plot, 
                         scene = scene, 
                         width = 600, height = 600)

  if(hide.colourbar==TRUE){
    plot <- plotly::hide_colorbar(plot)
  }

  if( output.png==TRUE){
    plotly::orca(plot, file =  file.name)
  }

  return(plot)

}

set.seed(1993)
volume <- sample(1:20, 14) #randomly generated data for 14 structures
make_ggseg3d(volume, 
             colour.pal = viridis::viridis(n=length(volume)), 
             output.png = T,
             file.name = "data/R/fig1e.png")

```

```{r Fig1F, eval=FALSE}
#Figure 1F (Cortical ROI using ggseg)

library(ggseg)
library(ggplot2)
#devtools::install_github("LCBC-UiO/ggsegGlasser")
library(ggsegGlasser)

#create dataframe with region names and hemisphere for the glasser atlas
base_atlas <- as.data.frame(na.omit(cbind(glasser$data$region, glasser$data$hemi)))

colnames(base_atlas) <- c("region", "hemi")

set.seed(1993) #set a random seed (good practice for reproducibility)
Effect <- rnorm(dim(base_atlas)[1]) #generate a random numbers for each roi in the atlas

base_atlas <- cbind(Effect, base_atlas)

#Cortex
cortex <- ggseg(atlas = glasser,
                .data = base_atlas,
                mapping=aes(fill=Effect), 
                position="stacked",
                colour="black",
                size=.2,
                show.legend = T,
                plot.background = "white") +
  scale_fill_viridis_b() + 
  theme_void() +
  theme(legend.position = "bottom",
        legend.text=element_text(size=15),
        legend.title  = element_blank())

ggsave(filename = 'data/R/fig1f.png', plot = cortex, device = 'png', bg = 'white')

```

```{r Fig1G, eval=FALSE}
#Figure 1G (Vertex using ciftiTools)

library(rgl)
library(ggplot2)

# Load the package and point to the Connectome Workbench
library(ciftiTools)

#Need connectome workbench to use ciftiTools - set path below
ciftiTools.setOption("wb_path", "/Applications/workbench/")

#Define CIFTI file names for time series and surface files. 
cifti_fname <- ciftiTools::ciftiTools.files()$cifti["dtseries"]
surfL_fname <- ciftiTools.files()$surf["left"]
surfR_fname <- ciftiTools.files()$surf["right"]

# Read in CIFTI file
cii <- read_cifti(
  cifti_fname, brainstructures="all", 
  surfL_fname=surfL_fname,
  surfR_fname=surfR_fname)

#Plot mean time series on CIFTI surface
view_xifti_surface(cii, 
                   hemisphere = "both", 
                   view = "lateral", 
                   idx = 1,
                   colors = viridis::plasma(n = 100),
                   zlim = c(1.2,1.5),  
                   legend_embed = T)

rgl.snapshot("data/R/fig1g.png") #save output in RGL window as 
rgl.close()

```

```{r Fig1H, eval=FALSE}
#Figure 1H (Vertex using fsbrain)

library(fsbrain)
library(rgl)

#download example freesurfer data which comes with the fsbrain package
#download_optional_data() 
#download_fsaverage(accept_freesurfer_license = TRUE)

subjects_dir <- get_optional_data_filepath("subjects_dir")

subject_id <- 'subject1'

colourmap <- colorRampPalette(viridis::plasma(n = 1000)) #set colourmap

# set figure size (optional)
fsbrain.set.default.figsize(700, 700)

# generate figure
brain = vis.subject.morph.native(subjects_dir, 
                                 subject_id, 'sulc', 
                                 cortex_only = T, 
                                 views=NULL, 
                                 draw_colorbar = FALSE,
                                 makecmap_options = list('colFn'=colourmap))

img = export(brain, draw_colorbar = FALSE, output_img = "data/R/fig1h.png") #write out figure

```

```{r Fig1I, eval=FALSE}
#Fig1I (Edges visualized using brainconn)

#devtools::install_github("sidchop/brainconn")
library(brainconn)

conmat <- example_weighted_undirected #one of the example matrices included with brainconn
degree <- rowSums(conmat)[which(rowSums(conmat)!=0)]

fig1i <- brainconn(atlas ="schaefer300_n7", 
                   conmat=conmat, 
                   node.size = degree/2,
                   view="ortho", 
                   edge.color.weighted = T,
                   background.alpha = 0.6, 
                   edge.color = scale_edge_colour_viridis(option = "C"))

ggsave(filename = 'data/R/fig1i.png', 
       plot = fig1i, 
       device = 'png',
       bg = 'white',
       height = 7, 
       units = "in")

```

\singlespacing

```{r Fig1 compile, eval=FALSE}
# This code block complies all the panels of the figure into one, largely using the patchwork package.

library(patchwork)
library(ggplot2)

fig_name <- c('fig1a', 'fig1b', 'fig1c', 'fig1d', 'fig1e', 'fig1f', 'fig1g1', 'fig1g2', 'fig1g3','fig1h', 'fig1i')
figs <- list()

for (f in 1:length(fig_name)) { 
  knitr::plot_crop(paste0('data/R/',fig_name[f],'.png'))
  figs[[f]] <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/R/',fig_name[f],'.png')),
                                                             width=ggplot2::unit(1,"npc"),
                                                             height=ggplot2::unit(1,"npc")))
}

patchwork <- figs[[1]]/figs[[2]]|((figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[7]]) + 
  plot_layout(heights = c(0.8,0.8,1.5))|((figs[[3]]/figs[[4]]/figs[[10]]) + 
                                           plot_layout(heights = c(0.5,1,1)))

patchwork + plot_annotation(tag_levels = 'A',
                            title = 'Examples of brain visualizations made in R') &
  theme(plot.tag = element_text(face = 'bold'),
        plot.title = element_text(face="bold"))

```